Effectful Semantics for Substructural SSA

### 2023-11-21 pldi24/paper.tex
Tap

% \section{Proofs}

% \subsection{Weakening and Substitution}

% \begin{theorem}[Weakening] \
%   \begin{itemize}
%     \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
%     \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
%     \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then
%     \(\haslb{\Gamma}{t}{\ms{L}}\)
%     \item If \(\lwk{\ms{L}}{\ms{K}}\), \(\haslb{\Gamma}{t}{\ms{L}}\) then
%     \(\haslb{\Gamma}{t}{\ms{K}}\)
%   \end{itemize}
% \end{theorem}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Substitution Splitting] 
%   If \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   then there exist \(\Theta_\Delta\), \(\Theta_\Xi\) such that:
%   \begin{itemize}
%     \item \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\)
%     \item \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\)
%     \item \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} 
%   We proceed by induction on derivations \(\csplits{\Gamma}{\Delta}{\Xi}\). In
%   the \rle{split-nil} case, we have that \(\issubst{\gamma}{\cdot}{\cdot}\) by
%   \rle{subst-nil}. For all other cases, since \(\issubst{\gamma}{\Theta}{\Gamma,
%   \thyp{x}{A}{q}}\), we have by \rle{subst-cons} that
%   \(\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}\),
%   \(\issubst{\gamma}{\Theta_\Gamma}{\Gamma}\),
%   \(\hasty{\Theta_\Gamma}{1}{\gamma(x)}{A}{q}\). Since
%   \(\csplits{\Gamma}{\Delta}{\Xi}\), by induction, we have that
%   \(\csplits{\Theta_\Gamma}{\Theta_\Delta}{\Theta_\Xi}\) where \(\Theta_\Delta\)
%   and \(\Theta_\Xi\) have the desired properties above.
%   \begin{itemize}
%     \item \rle{split-left}: by left split association, we have that there exists
%     \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\csplits{\Theta}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_\Xi}\), \(\csplits{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_\Delta}{\Theta_x}\). Using the fact that
%     \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), we can conclude by
%     \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Delta, \thyp{x}{A}{q}}\); the desired result hence
%     follows.
%     \item \rle{split-right}: by right split association, we have that there
%     exists \(\Theta_{\Xi, \thyp{x}{A}{q}}\),
%     \(\csplits{\Theta}{\Theta_\Delta}{\Theta_{\Xi, \thyp{x}{A}{q}}}\),
%     \(\csplits{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Theta_\Xi}{\Theta_x}\). Using the
%     fact that \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by
%     \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
%     \thyp{x}{A}{q}}\); the desired result hence follows.
%     \item \rle{split-dup}: since \(\tlin{A}{\trel}\), it follows by
%     \rle{subst-cons} that \(\tlin{\Theta_x}{\trel}\), and hence that
%     \(\csplits{\Theta_x}{\Theta_x}{\Theta_x}\). Therefore, by split association,
%     we have that there exist \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\Theta_{\Xi,
%     \thyp{x}{A}{q}}\) such that \(\csplits{\Theta}{\Theta_{\Delta,
%     \thyp{x}{A}{q}}}{\Theta_{\Xi, \thyp{x}{A}{q}}}\). Using the facts that
%     \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\),
%     \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by \rle{subst-cons}
%     that \(\issubst{\gamma}{\Theta_{\Delta, \thyp{x}{A}{q}}}{\Delta,
%     \thyp{x}{A}{q}}\), \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
%     \thyp{x}{A}{q}}\); the desired result hence follows.
%     \item \rle{split-drop}: since \(\tlin{A}{\taff}\), it follows by
%     \rle{subst-cons} that \(\tlin{\Theta_x}{\taff}\), and hence that
%     \(\cwk{\Theta_x}{\cdot}\). Therefore, by split association,
%     \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\), giving the desired result.
%   \end{itemize}
% \end{proof}

% \begin{lemma}[Substitution] 
%   Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   \begin{itemize}
%     \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\)
%     \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
%     \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K}, \ms{W}'\) such that 
%     \begin{itemize}
%       \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
%       \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
%       \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
%     \end{itemize}
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Label Substitution] 
%   Given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), 
%   \begin{itemize}
%     \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
%     \item For all \(\lhaslb{\ms{W}}{L}{\ms{L}}\), \(\lhaslb{\ms{W}}{[\mc{L}]L}{\ms{K}}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \subsection{Semantic Metatheory}

% \begin{lemma}[Semantic Upgrade]
%   For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
%   \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Semantic Weakening] 
%   If \(\cwk{\Gamma}{\Delta}\), then 
%   \begin{itemize}
%     \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
%       = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%     \]
%     \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
%     \[
%       \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
%       ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
%       = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
%     \]
%   \end{itemize}

%   Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
%   \[
%     \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%     ; \dnt{\lwk{\ms{L}}{\ms{K}}}
%     = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
%   \]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Substitution Coherence] \
%   \begin{itemize}
%     \item Given any two derivations \(D_1: \issubst{\gamma}{\Theta}{\Gamma}\), \(D_2: \issubst{\gamma'}{\Theta}{\Gamma}\) such that \(\forall x \in \Theta, \gamma(x) = \gamma'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
%     \item Given any two derivations \(D_1: \lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \lbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
%   \end{itemize}
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{lemma}[Semantic Substitution Splitting] 
%   Given \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Delta}{\Xi}\),
%   if \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\) and \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), then
%   \[
%     \dnt{\issubst{\gamma}{\Theta}{\Gamma}};\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
%     = \dnt{\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}};
%       \dnt{\issubst{\gamma}{\Theta_\Delta}{\Delta}}
%       \otimes \dnt{\issubst{\gamma}{\Theta_\Xi}{\Xi}}
%   \]
% \end{lemma}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

% \begin{theorem}[Semantic Substitution] 
%   Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
%   \begin{itemize}
%     \item For all \(\hasty{\Gamma}{p}{a}{A}{q}\), 
%     \[
%       \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
%       ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
%       = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}
%     \]
%     \item For all 
%       \(\haslb{\Gamma}{t}{\ms{L}}\), 
%       \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\), 
%       \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
%     \[
%       \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
%       ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
%       = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
%       ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
%     \]
%     \item For all 
%       \(\lhaslb{\ms{W}}{L}{\ms{L}}\), 
%       \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
%       \(\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}\),
%       \[
%         \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
%         ; \dnt{\lhaslb{\ms{W'}}{[\gamma]L}{\ms{K}}}
%         = \dnt{\lhaslb{\ms{W}}{L}{\ms{L}}}
%         ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}}
%         + \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
%       \]
%   \end{itemize}
%   Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\),
%   \begin{itemize} 
%     \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), we have
%     \[
%       \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
%       = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
%     \]
%     \item For all \(\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}\), we have
%     \[
%       \dnt{\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
%       + \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}}
%       = \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}};\dnt{\lhaslb{\ms{W}_{\ms{K}}}{[\mc{L}]L}{\ms{K}}}
%     \]
%     where all variables in \(\ms{W}\) not in \(\ms{L}\) are fresh w.r.t \(\mc{L}\).
%   \end{itemize}
% \end{theorem}

% \begin{proof} \
%   \TODO{this}
% \end{proof}

>>>>

\section{Proofs}

\begin{theorem}[Weakening] \
  \begin{enumerate}
    \item If \(\cwk{\Gamma}{\Delta}\) and \(q \subseteq r\),
    \(\hasty{\Delta}{p}{a}{A}{r}\), then \(\hasty{\Gamma}{p}{a}{A}{q}\)
    \item If \(\cwk{\Gamma}{\Delta}\), \(\haslb{\Delta}{t}{\ms{L}}\), then
    \(\haslb{\Gamma}{t}{\ms{L}}\)
    \item If \(\lwk{\ms{L}}{\ms{K}}\), \(\haslb{\Gamma}{t}{\ms{L}}\) then
    \(\haslb{\Gamma}{t}{\ms{K}}\)
  \end{enumerate}
\end{theorem}

\begin{proof}
  By induction: 
  \begin{enumerate}
    \item
    \item
    \item 
  \end{enumerate}
\end{proof}

\begin{lemma}[Substitution Splitting] 
  If \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Theta}{\Gamma}\),
  then there exist \(\Theta_\Delta\), \(\Theta_\Xi\) such that:
  \begin{itemize}
    \item \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\)
    \item \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\)
    \item \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\)
  \end{itemize}
\end{lemma}

\begin{proof} 
  We proceed by induction on derivations \(\csplits{\Gamma}{\Delta}{\Xi}\). In
  the \rle{split-nil} case, we have that \(\issubst{\gamma}{\cdot}{\cdot}\) by
  \rle{subst-nil}. For all other cases, since \(\issubst{\gamma}{\Theta}{\Gamma,
  \thyp{x}{A}{q}}\), we have by \rle{subst-cons} that
  \(\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}\),
  \(\issubst{\gamma}{\Theta_\Gamma}{\Gamma}\),
  \(\hasty{\Theta_\Gamma}{1}{\gamma(x)}{A}{q}\). Since
  \(\csplits{\Gamma}{\Delta}{\Xi}\), by induction, we have that
  \(\csplits{\Theta_\Gamma}{\Theta_\Delta}{\Theta_\Xi}\) where \(\Theta_\Delta\)
  and \(\Theta_\Xi\) have the desired properties above.
  \begin{itemize}
    \item \rle{split-left}: by left split association, we have that there exists
    \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\csplits{\Theta}{\Theta_{\Delta,
    \thyp{x}{A}{q}}}{\Theta_\Xi}\), \(\csplits{\Theta_{\Delta,
    \thyp{x}{A}{q}}}{\Theta_\Delta}{\Theta_x}\). Using the fact that
    \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), we can conclude by
    \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Delta,
    \thyp{x}{A}{q}}}{\Delta, \thyp{x}{A}{q}}\); the desired result hence
    follows.
    \item \rle{split-right}: by right split association, we have that there
    exists \(\Theta_{\Xi, \thyp{x}{A}{q}}\),
    \(\csplits{\Theta}{\Theta_\Delta}{\Theta_{\Xi, \thyp{x}{A}{q}}}\),
    \(\csplits{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Theta_\Xi}{\Theta_x}\). Using the
    fact that \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by
    \rle{subst-cons} that \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
    \thyp{x}{A}{q}}\); the desired result hence follows.
    \item \rle{split-dup}: since \(\tlin{A}{\trel}\), it follows by
    \rle{subst-cons} that \(\tlin{\Theta_x}{\trel}\), and hence that
    \(\csplits{\Theta_x}{\Theta_x}{\Theta_x}\). Therefore, by split association,
    we have that there exist \(\Theta_{\Delta, \thyp{x}{A}{q}}\), \(\Theta_{\Xi,
    \thyp{x}{A}{q}}\) such that \(\csplits{\Theta}{\Theta_{\Delta,
    \thyp{x}{A}{q}}}{\Theta_{\Xi, \thyp{x}{A}{q}}}\). Using the facts that
    \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\),
    \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), we can conclude by \rle{subst-cons}
    that \(\issubst{\gamma}{\Theta_{\Delta, \thyp{x}{A}{q}}}{\Delta,
    \thyp{x}{A}{q}}\), \(\issubst{\gamma}{\Theta_{\Xi, \thyp{x}{A}{q}}}{\Xi,
    \thyp{x}{A}{q}}\); the desired result hence follows.
    \item \rle{split-drop}: since \(\tlin{A}{\taff}\), it follows by
    \rle{subst-cons} that \(\tlin{\Theta_x}{\taff}\), and hence that
    \(\cwk{\Theta_x}{\cdot}\). Therefore, by split association,
    \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\), giving the desired result.
  \end{itemize}
\end{proof}

\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K}, \ms{W}'\) such that 
    \begin{itemize}
      \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
      \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
      \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
    \end{itemize}
  \end{itemize}
\end{lemma}

\begin{proof} \
  \TODO{this}
\end{proof}

\begin{lemma}[Label Substitution] 
  Given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), 
  \begin{itemize}
    \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
    \item For all \(\lhaslb{\ms{W}}{L}{\ms{L}}\), \(\lhaslb{\ms{W}}{[\mc{L}]L}{\ms{K}}\)
  \end{itemize}
\end{lemma}

\begin{proof} \
  \TODO{this}
\end{proof}

\begin{lemma}[Semantic Upgrade]
  For all \(\Gamma, a, A\), if \(\hasty{\Gamma}{1}{a}{A}{q}\), then
  \[\dnt{\hasty{\Gamma}{0}{a}{A}{q}} = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{}\]
\end{lemma}

\begin{proof} \
  \TODO{this}
\end{proof}

\begin{lemma}[Semantic Weakening] 
  If \(\cwk{\Gamma}{\Delta}\), then 
  \begin{itemize}
    \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
    \[
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
      = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
    \]
    \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
    \[
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
      = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
    \]
  \end{itemize}

  Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
  \[
    \dnt{\haslb{\Gamma}{t}{\ms{L}}}
    ; \dnt{\lwk{\ms{L}}{\ms{K}}}
    = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
  \]
\end{lemma}

\begin{proof} \
  \TODO{this}
\end{proof}

\begin{lemma}[Substitution Coherence] \
  \begin{itemize}
    \item Given any two derivations \(D_1: \issubst{\gamma}{\Theta}{\Gamma}\), \(D_2: \issubst{\gamma'}{\Theta}{\Gamma}\) such that \(\forall x \in \Theta, \gamma(x) = \gamma'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
    \item Given any two derivations \(D_1: \lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \lbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
  \end{itemize}
\end{lemma}

\begin{proof} \
  \TODO{this}
\end{proof}

\begin{lemma}[Semantic Substitution Splitting] 
  Given \(\csplits{\Gamma}{\Delta}{\Xi}\) and \(\issubst{\gamma}{\Delta}{\Xi}\),
  if \(\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}\) and \(\issubst{\gamma}{\Theta_\Delta}{\Delta}\), \(\issubst{\gamma}{\Theta_\Xi}{\Xi}\), then
  \[
    \dnt{\issubst{\gamma}{\Theta}{\Gamma}};\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
    = \dnt{\csplits{\Theta}{\Theta_\Delta}{\Theta_\Xi}};
      \dnt{\issubst{\gamma}{\Theta_\Delta}{\Delta}}
      \otimes \dnt{\issubst{\gamma}{\Theta_\Xi}{\Xi}}
  \]
\end{lemma}

\begin{proof} \
  \TODO{this}
\end{proof}

\begin{theorem}[Semantic Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item For all \(\hasty{\Gamma}{p}{a}{A}{q}\), 
    \[
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
      ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
      = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}
    \]
    \item For all 
      \(\haslb{\Gamma}{t}{\ms{L}}\), 
      \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\), 
      \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
    \[
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
      ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
      = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
      ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
    \]
    \item For all 
      \(\lhaslb{\ms{W}}{L}{\ms{L}}\), 
      \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
      \(\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}\),
      \[
        \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
        ; \dnt{\lhaslb{\ms{W'}}{[\gamma]L}{\ms{K}}}
        = \dnt{\lhaslb{\ms{W}}{L}{\ms{L}}}
        ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}}
        + \dnt{\lbsubst{\substlbs{\gamma}{\ms{W'}}}{\ms{W}}{\ms{W'}}}
      \]
  \end{itemize}
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\),
  \begin{itemize} 
    \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), we have
    \[
      \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
    \]
    \item For all \(\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}\), we have
    \[
      \dnt{\lhaslb{\ms{W}_{\ms{L}}}{L}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      + \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}}
      = \dnt{\lbsubst{\mc{L}}{\ms{W}_{\ms{L}}}{\ms{W}_{\ms{K}}}};\dnt{\lhaslb{\ms{W}_{\ms{K}}}{[\mc{L}]L}{\ms{K}}}
    \]
    where all variables in \(\ms{W}\) not in \(\ms{L}\) are fresh w.r.t \(\mc{L}\).
  \end{itemize}
\end{theorem}

\begin{proof} \
  \TODO{this}
\end{proof}
